---
title: "Case Study 2 - EDA"
author: "Nikhil Gupta"
date: "`r Sys.time()`"
always_allow_html: yes
output:
  github_document:
    toc: true
    toc_depth: 6
  html_document:
    toc: true
    toc_float: true
    toc_depth: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
version
```

** From text (Chapter 2)**

# Section 2.2

```{r}
menTables = readRDS("../../data/CBMenTextTables.rds")
menTables[['2012']][1:10]
```

```{r}
menTables[['2006']][1:10]
```

```{r}
els = menTables[['2012']]
```


```{r}
eqIndex = grep("^===", els)
eqIndex
```

```{r}
spacerRow = els[eqIndex]
headerRow = els[eqIndex - 1] %>% tolower()
body = els[ -(1:eqIndex) ]
```


```{r}
ageStart = regexpr("ag", headerRow)
```

```{r}
age = substr(body, start = ageStart, stop = ageStart + 1)
head(age)
```

```{r}
summary(as.numeric(age))
```

```{r}
# Here the g in gregexpr() stands for “global,” which means that the function searches for
```


```{r}
# multiple matches in the string, not just the first match.
blankLocs = gregexpr(" ", spacerRow)
blankLocs
```


```{r}
searchLocs = c(0, blankLocs[[1]])
searchLocs
```


```{r}
Values = mapply(
  substr,
  list(body),
  start = searchLocs[-length(searchLocs)] + 1,
  stop = searchLocs[-1] - 1
  )
Values[1:10,]
```


```{r}
findColLocs = function(spacerRow){
  # In the function, we safeguard against the last character in the row of '=' characters not being a blank, we add
  # an additional element to the end of the vector of locations that is one character more than the length of the string.
  spaceLocs = gregexpr(" ", spacerRow)[[1]]
  rowLength = nchar(spacerRow)
  if(substring(spacerRow, rowLength, rowLength) != " "){
    return(c(0, spaceLocs, rowLength + 1))
  }
  else{
    return(c(0, spaceLocs))
  }
}
```

```{r}
#' @description Function to determine the start and end location of a Column
#' @returns The Start and End Position of the Column
selectCols = function(colNames, headerRow, searchLocs){
  sapply(
    colNames,
    function(name, headerRow, searchLocs){
      startPos = regexpr(name, headerRow)[[1]]
      if (startPos == -1){
        return(c(NA, NA))
      }
      index = sum(startPos >= searchLocs)
      
      # We can easily solve both of these problems by including the value in the “blank” space between columns (see text book).
      # return(c(searchLocs[index] + 1, searchLocs[index + 1] - 1))
      return(c(searchLocs[index] + 1, searchLocs[index + 1]))
    },
    headerRow = headerRow, searchLocs = searchLocs
  )
}
```

```{r}
searchLocs = findColLocs(spacerRow)
ageLoc = selectCols("ag", headerRow, searchLocs)
ages = mapply(
  substr,
  list(body),
  start = ageLoc[1,], stop = ageLoc[2,]
)
summary(as.numeric(ages))
```


```{r}
# Since the column names vary somewhat from year to year, we use only the first few
# characters that uniquely identify the desired columns, e.g.,
shortColNames = c("name", "home", "ag", "gun", "net", "time")
```


```{r}
# Also, if a file does not have one of the desired variables, then we want the values for that variable to be NA.
locCols = selectCols(shortColNames, headerRow, searchLocs)
Values = mapply(
  substr,
  list(body),
  start = locCols[1, ],
  stop = locCols[2, ]
)
colnames(Values) = shortColNames
head(Values)
```

```{r}
tail(Values)
```

```{r}
extractVariables = function(file, varNames =c("name", "home", "ag", "gun", "net", "time")){
  # Find the index of the row with =s
  eqIndex = grep("^===", file)
  
  # Extract the two key rows and the data
  spacerRow = file[eqIndex]
  headerRow = tolower(file[ eqIndex - 1 ])
  body = file[ -(1 : eqIndex) ]
  
  # Remove Comments (Starting with optional space and #)
  comments = grep("^[[:space:]]*#", body)
  if(length(comments) > 0){
    body = body[-comments]
  }
  
  # Obtain the starting and ending positions of variables
  searchLocs = findColLocs(spacerRow)
  locCols = selectCols(varNames, headerRow, searchLocs)
  
  # Extract the values of these variables
  Values = mapply(
    substr, list(body),
    start = locCols[1, ],
    stop = locCols[2, ]
  )
  colnames(Values) = varNames
  
  # Remove empty rows
  blanks = grep("^[[:blank:]]*$", Values)
  if(length(blanks) > 0){
    Values = Values[-blanks, ]
  }
  
  return(invisible(Values))
}
```


```{r}
menResMat = lapply(menTables, extractVariables)
length(menResMat)
```

```{r}
sapply(menResMat, nrow)
```

# Section 2.3

## Cleaning Age

```{r}
age = as.numeric(menResMat[['2012']][ , 'ag'])
tail(age)
```

```{r}
age = sapply(
  menResMat,
  function(x) as.numeric(x[ , 'ag'])
)
```

```{r}
boxplot(age, ylab = "Age", xlab = "Year")
```

We see that in 2003, the age values are shifted to the right one space in comparison to the
location of the '=' characters. This means that we are picking up only the digit in the tens
place. In 2006, some but not all of the rows have values that are off by one character.

```{r}
head(menTables[['2003']])
```

```{r}
menTables[['2006']][2200:2205]
```

```{r}
head(menTables[['2009']],10)
```

```{r}
sapply(age, function(x) sum(is.na(x)))
```


```{r}
age2001 = age[["2001"]]
```


```{r}
# grep("^===", menTables[['2001']])
```


```{r}
# badAgeIndex = which(is.na(age2001)) + 5
# badAgeIndex
# menTables[['2001']][badAgeIndex]
```


```{r}
# blanks = grep("^[[:blank:]]*$", menTables[['2001']])
# blanks
```


```{r}
menResMat[['2001']][age2001 < 5,]
```

## Cleaning Time

```{r}
charTime = menResMat[['2012']][, 'time']
head(charTime, 5)
tail(charTime, 5)
```

```{r}
timePieces = strsplit(charTime, ":")
timePieces[[1]]
tail(timePieces, 1)
```

```{r}
timePieces = sapply(timePieces, as.numeric)
runTime = sapply(
  timePieces,
  function(x) {
    if (length(x) == 2) x[1] + x[2]/60
    else 60*x[1] + x[2] + x[3]/60
  }
)
summary(runTime)
```


```{r}
#' @description Converts character time columns to calculated numeric values
convertTime = function(charTimeValues){
  
  timePieces = charTimeValues %>% 
    strsplit(":") %>%
    sapply(as.numeric)
  
  runTime = sapply(
    timePieces,
    function(x) {
      if (length(x) == 2) x[1] + x[2]/60
      else 60*x[1] + x[2] + x[3]/60
    }
  )
  return(runTime)
}
```


```{r}
summary(convertTime(charTime))
```

## DataFrame

```{r}
createDF = function(Res, year, sex){
  # Determine which time to use
  if(!is.na(Res[1, 'net'])){
    useTime = Res[ , 'net']
  }
  else if(!is.na(Res[1, 'gun'])){
    useTime = Res[ , 'gun']
  }
  else{
    useTime = Res[ , 'time']
  }
  
  # Remove # and * and blanks from time
  useTime = gsub("[#\\*[:blank:]]", "", useTime)
  # Drop rows with no time
  Res = Res[ useTime != "", ]
  
  # Convert from hh:mm:ss to single numeric value
  runTime = convertTime(useTime[useTime != ""])
  
  Results = data.frame(
    year = rep(year, nrow(Res)),
    sex = rep(sex, nrow(Res)),
    name = Res[ , 'name'],
    home = Res[ , 'home'],
    age = as.numeric(Res[, 'ag']),
    runTime = runTime,
    stringsAsFactors = FALSE
  )
  return(invisible(Results))
}
```

```{r}
menDF = mapply(
  createDF,
  menResMat,
  year = 1999:2012,
  sex = rep("M", length(1999:2012)),
  SIMPLIFY = FALSE
)
```

```{r}
# It is likely that the conversion problems are coming from the conversion of time from a
# character string into minutes because we have already handled the conversion of age.
sapply(menDF, function(x) sum(is.na(x$runTime)))
```

We find that these are caused by runners who completed half the race but have no final times and by runners who have a footnote after their time,

```{r}
menDF[['2001']][is.na(menDF[['2001']]$runTime), ]
menDF[['2002']][is.na(menDF[['2002']]$runTime), ]
```

```{r}
head(menDF[['2006']])
```

```{r}
# Hometown and Net Time are Combined
menTables[['2006']][1:20]
```

# TODO: Fix 2006 Results (Header)

```{r}
cbMen = do.call(rbind, menDF)
```

```{r}
dim(cbMen)
```

```{r}
saveRDS(cbMen, file = "../../data/cbMen.rds")
```

# Section 2.4: Exploring Run Times (Male Runners)

```{r}
plot(
  runTime ~ age, data = cbMen, ylim = c(40, 180),
  xlab = "Age (years)", ylab = "Run Time (minutes)"
)
```

**TODO**: Convert to GGPLOT with faceting or color coding by year

## 2.4.1 Making Plots with Many Observations

```{r}
library(RColorBrewer)
ls("package:RColorBrewer")
```

```{r fig.width=8, fig.height=8}
display.brewer.all()
```


```{r}
Purples8 = brewer.pal(9, "Purples")[8]
Purples8
```

```{r}
# Adding Transparency
Purples8A = paste(Purples8, "14", sep = "")
```


**TODO**: Plot with Color (left as exercise to user)

```{r}
# Better Alternate
smoothScatter(
  y = cbMen$runTime, x = cbMen$age,
  ylim = c(40, 165), xlim = c(15, 85),
  xlab = "Age (years)", ylab = "Run Time (minutes)"
)
```


```{r}
# Graphing Sumary Statistics
cbMenSub = cbMen[cbMen$runTime > 30 & !is.na(cbMen$age) & cbMen$age > 15, ]
ageCat = cut(cbMenSub$age, breaks = c(seq(15, 75, 10), 90))
table(ageCat)
```

```{r}
plot(
  cbMenSub$runTime ~ ageCat,
  xlab = "Age (years)", ylab = "Run Time (minutes)"
)
```

## 2.4.2: Fitting Models to Average Performance

```{r}
# TODO: When 2006 is fixed, this should include all the data.
# Right now, it excludes 2006 data
dim(cbMenSub)
cbMenSub = cbMenSub[which(!is.na(cbMenSub$runTime)),]
dim(cbMenSub)
```

```{r}
lmAge = lm(runTime ~ age, data = cbMenSub)
summary(lmAge)
```

```{r}
# TODO: Fails since there are NA Values
smoothScatter(
  x = cbMenSub$age, y = lmAge$residuals,
  xlab = "Age (years)", ylab = "Residuals"
)
abline(h = 0, col = "purple", lwd = 3)
```


```{r}
resid.lo = loess(
  resids ~ age,
  data = data.frame(
    resids = residuals(lmAge),
    age = cbMenSub$age
  )
)
```

```{r}
age20to80 = 20:80
resid.lo.pr = predict(resid.lo, newdata = data.frame(age = age20to80))
```

```{r}
smoothScatter(
  x = cbMenSub$age, y = lmAge$residuals,
  xlab = "Age (years)", ylab = "Residuals"
)
abline(h = 0, col = "purple", lwd = 3)
lines(x = age20to80, y = resid.lo.pr, col = "green", lwd = 3, lty = 2)
```

### LOESS Model
```{r}
menRes.lo = loess(runTime ~ age, cbMenSub)
menRes.lo.pr = predict(menRes.lo, data.frame(age = age20to80))
```

```{r}
plot(x = age20to80, y = menRes.lo.pr, col = "green", lwd = 3, lty = 2)
```

### Piecewise Linear Model

```{r}
over50 = pmax(0, cbMenSub$age - 50)
lmOver50 = lm(runTime ~ age + over50, data = cbMenSub)
summary(lmOver50)
```


```{r}
decades = seq(30, 60, by = 10)
overAge = lapply(decades,
function(x) pmax(0, (cbMenSub$age - x)))
names(overAge) = paste("over", decades, sep = "")
overAge = as.data.frame(overAge)
tail(overAge)
```


```{r}
lmPiecewise = lm(
  runTime ~ . ,
  data = cbind(cbMenSub[, c("runTime", "age")], overAge)
)
summary(lmPiecewise)
```

```{r}
overAge20 = lapply(decades, function(x) pmax(0, (age20to80 - x)))
names(overAge20) = paste("over", decades, sep = "")
overAgeDF = cbind(age = data.frame(age = age20to80), overAge20)
tail(overAgeDF)
```


```{r}
predPiecewise = predict(lmPiecewise, overAgeDF)
```


```{r}
plot(
  predPiecewise ~ age20to80,
  type = "l", col = "purple", lwd = 3,
  xlab = "Age (years)", ylab = "Run Time Prediction"
)
lines(x = age20to80, y = menRes.lo.pr, col = "green", lty = 2, lwd = 3) # LOESS
legend(
  "topleft", col = c("purple", "green"),
  lty = c(1, 2), lwd= 3,
  legend = c("Piecewise Linear", "Loess Curve"),
  bty = "n"
)
```

## 2.4.3 Cross Sectional Data and Covariates

* Issue 1: Comparison across ages: Younger might be professional runners, older mighy be amateurs
* Issue 2: Composition of ages might have changed over the years

```{r}
summary(cbMenSub$runTime[cbMenSub$year == 1999])
summary(cbMenSub$runTime[cbMenSub$year == 2012])
```

```{r}
age1999 = cbMenSub[ cbMenSub$year == 1999, "age" ]
age2012 = cbMenSub[ cbMenSub$year == 2012, "age" ]
```

```{r}
plot(
  density(age1999, na.rm = TRUE),
  ylim = c(0, 0.05), col = "purple",
  lwd = 3, xlab = "Age (years)", main = ""
)
lines(
  density(age2012, na.rm = TRUE),
  lwd = 3, lty = 2, col="green"
)
legend("topleft", col = c("purple", "green"), lty= 1:2, lwd = 3,
legend = c("1999", "2012"), bty = "n")
```


```{r}
mR.lo99 = loess(runTime ~ age, cbMenSub[ cbMenSub$year == 1999,])
mR.lo.pr99 = predict(mR.lo99, data.frame(age = age20to80))
mR.lo12 = loess(runTime ~ age, cbMenSub[ cbMenSub$year == 2012,])
mR.lo.pr12 = predict(mR.lo12, data.frame(age = age20to80))
plot(
  mR.lo.pr99 ~ age20to80,
  type = "l", col = "purple", lwd = 3,
  xlab = "Age (years)", ylab = "Fitted Run Time (minutes)"
)
lines(
  x = age20to80, y = mR.lo.pr12,
  col = "green", lty = 2, lwd = 3
)
legend("topleft", col = c("purple", "green"), lty = 1:2, lwd = 3,
legend = c("1999", "2012"), bty = "n")
```

# Section 2.5

##Constructing a Record for an Individual Runner across Years

```{r}
```


```{r}
```


# Section 2.6

## Modeling the Change in Running for Individuals

```{r}
```


```{r}
```

