---
title: "Case Study 2"
author: "Nikhil, Moro, Bhuvana"
date: "`r Sys.time()`"
output:
  word_document:
    toc: yes
    toc_depth: '6'
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
  github_document:
    toc: yes
    toc_depth: 6
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, plotly, naniar)
```

# reading Data from ETL


```{r cars}
data = readRDS(file='../../data/data.rds')
dataRaw = readRDS(file='../../data/dataRaw.rds')
```

```{r}
head(data)
```

```{r}
message("Checking list of ages")
sort(unique(data$Age))
message("Checking list of Divisions")
table(data$Division)
```

# Missing Values

Analyzing the rows with **NR* division

```{r}
data %>% 
  dplyr::filter(Division == "NR")
```
Analyzing the rows with null Age

```{r}
data %>% 
  dplyr::filter(is.na(Age))
```

Analyzing the rows with null PiD

```{r}
data %>% 
  dplyr::filter(is.na(PiD))
```
Analyzing the rows with NA values in the Hometown, State, and country

```{r}
data %>% 
  dplyr::filter(is.na(Hometown) & is.na(HomeState) & is.na(HomeCountry))
```
Analyzing the rows with NA values in the Home State and Country

```{r}
data %>% 
  dplyr::filter(is.na(HomeState) & is.na(HomeCountry))
```
Analyzing the rows with NA values in the Home Country

```{r}
data %>% 
  dplyr::filter(is.na(HomeCountry))
```
Checking Home Countries

```{r}
sort(table(data$HomeCountry))
```
Some Home Countries have incorrenct naming
```{r}
table(dataRaw[data$HomeCountry == 'AE',]$Hometown)
table(dataRaw[data$HomeCountry == 'AP',]$Hometown)
table(dataRaw[data$HomeCountry == 'SK',]$Hometown)
table(dataRaw[data$HomeCountry == 'VI',]$Hometown)
table(dataRaw[data$HomeCountry == 'GBR',]$Hometown)
table(dataRaw[data$HomeCountry == 'PR',]$Hometown)
table(dataRaw[data$HomeCountry == 'ON',]$Hometown)
```

Fixing incorrect naming when possible
```{r}
data = data %>%
  mutate(HomeState = ifelse(HomeCountry =='VI','VA',HomeState)
         ,HomeCountry = recode(HomeCountry
                               , 'GBR' = 'GREAT BRITAIN'
                               , 'PR' = 'PUERTO RICO'
                               , 'ON' = 'CANADA' #Toronto
                               , 'VI' = 'USA' #Virginia
         ))
```
  
  
Converting Division with NR to null

```{r}
data = data %>%
  naniar::replace_with_na(
    replace = list(
      Division = "NR"
    )
  )
```


Counting NA per each field

```{r}
data %>% 
  select_if(function(x) any(is.na(x))) %>% 
  summarise_each(~ sum(is.na(.))) 
```

Plotting NA records

```{r fig.height=6, fig.width=7}
naniar::gg_miss_upset(data,nsets = ncol(data))
#naniar::vis_miss(data, warn_large_data=F, sort_miss=T)
```

#Cleaning of dataset

removing records with NA in TDivision, Age, PId, TID

```{r}
removeRecs = with(data, is.na(Division) | is.na(Pace) | is.na(PiS) |
                    is.na(PiD) | is.na(TiS) | is.na(TiD) | is.na(Age))

message("Removing ", sum(removeRecs), " invalid Records")
data = data[! removeRecs,]
```


# Basic EDA
```{r}
plotdata = data %>% 
  group_by(year) %>% 
  summarise(count=n()) 

p = plotdata %>% 
  ggplot(aes(x=year, y=count)) +
  geom_line() +
  geom_point()
print(p)
#ggplotly(p, tooltip="text")

p = plotdata %>% 
  ggplot(aes(x = year, y = count)) + 
  geom_bar(stat = "identity")
print(p)
#ggplotly(p, tooltip="text")
```

```{r}
participants_year_div = data %>% 
  group_by(year, Division) %>% 
  summarise(num_participants = n())
```

```{r fig.height=5, fig.width=8}
participants_year_div %>% 
  ggplot(aes(x=Division, y=year, fill=num_participants)) +
  geom_tile() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_gradient2(low = "green", mid='darkorange', high = "darkred", na.value = NA
                       #,midpoint=mean(c(max(participants_year_div$num_participants),min(participants_year_div$num_participants)))) + 
                       ,midpoint=quantile(participants_year_div$num_participants, 0.9)[[1]]) + 
  geom_text(data = participants_year_div, label = participants_year_div$num_participants)

```

```{r}
# Alternate colored by Division
plotdata = data %>% 
  group_by(year, Division) %>% 
  summarise(count = n())

p = plotdata %>% 
  ggplot(aes(x = year, y = count, fill = Division)) + 
  geom_bar(stat = "identity", position = "stack") 
print(p)
#ggplotly(p, tooltip="text")
```



# Business Analysis

## Age Brackets by Year

```{r fig.height=7, fig.width=10}
plotdata_by_year = data %>% 
  group_by(year) %>% 
  summarise(count_year = n())

plotdata_by_year_div = data %>% 
  group_by(year, Division) %>% 
  summarise(count_year_div = n())

plotdata = plotdata_by_year %>% 
  plyr::join(plotdata_by_year_div, by = "year", type = "full") %>% 
  mutate(percent = round(count_year_div/count_year*100,1))

plotdata %>% 
  ggplot(aes(x = year, y = percent, fill =forcats::fct_rev( Division))) + 
  geom_bar(stat = "identity", position = "stack") +
  coord_flip() + 
  labs(fill='Division') + 
  geom_text(data=filter(plotdata,percent>2.5),aes(label = paste0(scales::percent(percent/100,accuracy=1)))
            , position = ggplot2::position_stack(vjust=0.5)
            , check_overlap = T, size=3,vjust=-0.5) + 
  geom_text(data=filter(plotdata,percent>2.5),aes(label = Division)
            , position = ggplot2::position_stack(vjust=0.5)
            , check_overlap = T, size=2.5,vjust=1)  + 
  scale_x_continuous(breaks = unique(plotdata$year))


```

## Race Times

```{r fig.height=4, fig.width=8}
p = data %>% 
  mutate_at("year", as.factor) %>% 
  ggplot(aes(x=year, y=TimeMins, fill=year)) + 
  geom_boxplot() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  coord_flip()
print(p)
#ggplotly(p, tooltip="text")
```


```{r fig.height=4, fig.width=8}
p = data %>% 
  # dplyr::filter(!(Division %in% c("W8099", "W7579"))) %>% 
  mutate_at("year", as.factor) %>% 
  ggplot(aes(x=Division, y=TimeMins, fill=Division)) + 
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  coord_flip()

print(p)
#ggplotly(p, tooltip="text")
```

```{r fig.height=12, fig.width=12}
p = data %>% 
  dplyr::filter(!(Division %in% c("W8099", "W7579", "W7074", "W6569"))) %>% 
  mutate_at("year", as.factor) %>% 
  ggplot(aes(x=year, y=TimeMins, fill=year)) + 
  geom_boxplot() +
  facet_wrap(. ~ Division, ncol=2) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  coord_flip()

print(p)
#ggplotly(p, tooltip="text")
```
