---
title: "Max_CS1"
author: "Max Moro"
date: "8/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
dataLoc = '../../data/'
library(tidyverse)
library(ggplot2)
```

# Read Data
```{r}
offline= readRDS(file.path(dataLoc,'offline.rds'))
subMacs = names(sort(table(offline$mac), decreasing = TRUE))[1:7]

online= readRDS(file.path(dataLoc,'online.rds'))

```

# Exploring

## Signal vs. Angle

```{r}
library(ggplot2)

p=ggplot(data=filter(offline, mac != "00:0f:a3:39:dd:cd")
         ,aes(y=signal,x=factor(angle))) +
  ggplot2::geom_violin()+
  ggplot2::facet_wrap(vars(mac))
p

```
## Density Signal by MAC Angle

```{r fig.height=10, fig.width=10}
p=ggplot(data=filter(offline, mac != "00:0f:a3:39:dd:cd")
         ,aes(x=signal) 
         ) +
  ggplot2::geom_density()+
  ggplot2::facet_wrap(vars(mac
                           ,paste0('Angle:',angle)))
p


```

## SD Signal Strenth by Mean 

```{r}
byLocAngleAP = with(offline,
                    by(offline, list(posXY, angle, mac),
                       function(x) x))

signalSummary =
  lapply(byLocAngleAP,
         function(oneLoc) {
           ans = oneLoc[1, ]
           ans$medSignal = median(oneLoc$signal)
           ans$avgSignal = mean(oneLoc$signal)
           ans$num = length(oneLoc$signal)
           ans$sdSignal = sd(oneLoc$signal)
           ans$iqrSignal = IQR(oneLoc$signal)
           ans
         })
offlineSummary = do.call("rbind", signalSummary)

offlineSummary[1:10]
```


```{r}
breaks = seq(-90, -30, by = 5)
p=ggplot(data=filter(offlineSummary, mac != "00:0f:a3:39:dd:cd")
         ,aes(y=sdSignal,x=cut(avgSignal,breaks=breaks)) 
         ) +
  ggplot2::geom_boxplot()

  
p
```
## Avg - Med Signal vs obs

```{r}

with(offlineSummary,
     smoothScatter((avgSignal - medSignal) ~ num,
                   xlab = "Number of Observations",
                   ylab = "mean - median"))
abline(h = 0, col = "#984ea3", lwd = 2)

lo.obj =
  with(offlineSummary,
       loess(diff ~ num,
             data = data.frame(diff = (avgSignal - medSignal),
                               num = num)))

lo.obj.pr = predict(lo.obj, newdata = data.frame(num = (70:120)))
lines(x = 70:120, y = lo.obj.pr, col = "#4daf4a", lwd = 2)
```

## Signal and Distnace

```{r fig.height=8, fig.width=8}

library(fields)
surfaceSS = function(data, sel_mac, sel_angle = 45) {
 
   oneAPAngle = filter(data, mac ==sel_mac & angle == sel_angle)
  x=data.frame(oneAPAngle$posX,oneAPAngle$posY)
  y=oneAPAngle$avgSignal
  smoothSS = Tps(x,y)
  vizSmooth = predictSurface(smoothSS)
  plot.surface(vizSmooth, type = "C")
  points(oneAPAngle$posX, oneAPAngle$posY, pch=19, cex = 0.5)
}

parCur = par(mfrow = c(2,2), mar = rep(1.5, 4))

mapply(surfaceSS, 
       sel_mac = subMacs[ rep(c(5, 1), each = 2) ], 
       sel_angle = rep(c(0, 135), 2),
       data = list(data = offlineSummary))

par(parCur)

```

## Signal Strenghts vs Distance

```{r fig.height=9, fig.width=10}

offlineSummary = subset(offlineSummary, mac != subMacs[2])

AP = matrix( c( 7.5, 6.3, 2.5, -.8, 12.8, -2.8,  
                1, 14, 33.5, 9.3,  33.5, 2.8),
            ncol = 2, byrow = TRUE,
            dimnames = list(subMacs[ -2 ], c("x", "y") ))

#AP

diffs = offlineSummary[,c("posX", "posY")] -  AP[ offlineSummary$mac, ]
offlineSummary$dist = sqrt(diffs[ , 1]^2 + diffs[ , 2]^2)
ggplot2::ggplot(data=offlineSummary,aes(x=dist, y=signal)) +
  geom_point(size=0.1) +
  facet_grid(rows=vars(mac),cols=vars(angle))
#xyplot(signal ~ dist | factor(mac) + factor(angle),        data = offlineSummary, pch = 19, cex = 0.3,       xlab ="distance")
```

# Nearest Neighbor 

## Prepare Data

```{r}
macs = unique(offlineSummary$mac)
#talling signal strengths
online = filter(online,mac %in% macs)
tabonlineXYA = table(online$posXY, online$angle)
tabonlineXYA[1:6, ]

```

### Change structure of online data

```{r}
keepVars = c("posXY", "posX","posY", "orientation", "angle")
byLoc = with(online, 
             by(online, list(posXY), 
                function(x) {
                  ans = x[1, keepVars]
                  avgSS = tapply(x$signal, x$mac, mean)
                  y = matrix(avgSS, nrow = 1, ncol = 6,dimnames = list(ans$posXY, names(avgSS)))
                  cbind(ans, y)
                }))

onlineSummary = do.call("rbind", byLoc)  

dim(onlineSummary)
names(onlineSummary)
```
## Select Orientationg

```{r}

m = 3; angleNewObs = 230
refs = seq(0, by = 45, length  = 8)
nearestAngle = roundOrientation(angleNewObs)

  
if (m %% 2 == 1) {
  angles = seq(-45 * (m - 1) /2, 45 * (m - 1) /2, length = m)
} else {
  m = m + 1
  angles = seq(-45 * (m - 1) /2, 45 * (m - 1) /2, length = m)
  if (sign(angleNewObs - nearestAngle) > -1) 
    angles = angles[ -1 ]
  else 
    angles = angles[ -m ]
}
angles = angles + nearestAngle
angles[angles < 0] = angles[ angles < 0 ] + 360
angles[angles > 360] = angles[ angles > 360 ] - 360

#select the obs within the angles
offlineSubset = offlineSummary[ offlineSummary$angle %in% angles, ]

#reaggregate
reshapeSS = function(data, varSignal = "signal", 
                     keepVars = c("posXY", "posX","posY")) {
  byLocation =
    with(data, by(data, list(posXY), 
                  function(x) {
                    ans = x[1, keepVars]
                    avgSS = tapply(x[ , varSignal ], x$mac, mean)
                    y = matrix(avgSS, nrow = 1, ncol = 6,
                               dimnames = list(ans$posXY,
                                               names(avgSS)))
                    cbind(ans, y)
                  }))

  newDataSS = do.call("rbind", byLocation)
  return(newDataSS)
}

trainSS = reshapeSS(offlineSubset, varSignal = "avgSignal")
```
### Functio to Select

```{r}


selectTrain = function(angleNewObs, signals = NULL, m = 1){
  # m is the number of angles to keep between 1 and 5
  refs = seq(0, by = 45, length  = 8)
  nearestAngle = roundOrientation(angleNewObs)
  
  if (m %% 2 == 1) 
    angles = seq(-45 * (m - 1) /2, 45 * (m - 1) /2, length = m)
  else {
    m = m + 1
    angles = seq(-45 * (m - 1) /2, 45 * (m - 1) /2, length = m)
    if (sign(angleNewObs - nearestAngle) > -1) 
      angles = angles[ -1 ]
    else 
      angles = angles[ -m ]
  }
  angles = angles + nearestAngle
  angles[angles < 0] = angles[ angles < 0 ] + 360
  angles[angles > 360] = angles[ angles > 360 ] - 360
  angles = sort(angles) 
  
  offlineSubset = signals[ signals$angle %in% angles, ]
  reshapeSS(offlineSubset, varSignal = "avgSignal")
}

train130 = selectTrain(130, offlineSummary, m = 3)
head(train130)
length(train130[[1]])

```
## Finding Nearest Neighbors

```{r}
findNN = function(newSignal, trainSubset) {
  diffs = apply(trainSubset[ , 4:9], 1, 
                function(x) x - newSignal)
  dists = apply(diffs, 2, function(x) sqrt(sum(x^2)) )
  closest = order(dists)
  return(trainSubset[closest, 1:3 ])
}
```

### Estimate XY

```{r}


predXY = function(newSignals, newAngles, trainData, 
                  numAngles = 1, k = 3){
  
  closeXY = list(length = nrow(newSignals))
  
  for (i in 1:nrow(newSignals)) {
    trainSS = selectTrain(newAngles[i], trainData, m = numAngles)
    closeXY[[i]] = 
      findNN(newSignal = as.numeric(newSignals[i, ]), trainSS)
  }

  estXY = lapply(closeXY, 
                 function(x) sapply(x[ , 2:3], 
                                    function(x) mean(x[1:k])))
  estXY = do.call("rbind", estXY)
  return(estXY)
}
estXYk3 = predXY(newSignals = onlineSummary[ , 6:11], 
                 newAngles = onlineSummary[ , 4], 
                 offlineSummary, numAngles = 3, k = 3)

estXYk1 = predXY(newSignals = onlineSummary[ , 6:11], 
                 newAngles = onlineSummary[ , 4], 
                 offlineSummary, numAngles = 3, k = 1)

```

### Error Plot

```{r}

floorErrorMap = function(estXY, actualXY, trainPoints = NULL, AP = NULL){
  
    plot(0, 0, xlim = c(0, 35), ylim = c(-3, 15), type = "n",
         xlab = "", ylab = "", axes = FALSE)
    box()
    if ( !is.null(AP) ) points(AP, pch = 15)
    if ( !is.null(trainPoints) )
      points(trainPoints, pch = 19, col="grey", cex = 0.6)
    
    points(x = actualXY[, 1], y = actualXY[, 2], 
           pch = 19, cex = 0.8 )
    points(x = estXY[, 1], y = estXY[, 2], 
           pch = 8, cex = 0.8 )
    segments(x0 = estXY[, 1], y0 = estXY[, 2],
             x1 = actualXY[, 1], y1 = actualXY[ , 2],
             lwd = 2, col = "red")
}


```



### Plot Error 

```{r}
#get tain point
trainPoints = offlineSummary[ offlineSummary$angle == 0 & 
                              offlineSummary$mac == "00:0f:a3:39:e1:c0" ,
                        c("posX", "posY")]
```

#### K=1
```{r}

floorErrorMap(estXYk3, onlineSummary[ , c("posX","posY")], 
              trainPoints = trainPoints, AP = AP)
```

#### K=3
```{r}

floorErrorMap(estXYk1, onlineSummary[ , c("posX","posY")], 
              trainPoints = trainPoints, AP = AP)
```

### Error Plot

```{r}

calcError = function(estXY, actualXY) sum( rowSums( (estXY - actualXY)^2) )

actualXY = onlineSummary[ , c("posX", "posY")]
sapply(list(estXYk1, estXYk3), calcError, actualXY)
```

## Cross Validation 


```{r}
#select and reshape
v = 11
permuteLocs = sample(unique(offlineSummary$posXY))
permuteLocs = matrix(permuteLocs, ncol = v, 
                     nrow = floor(length(permuteLocs)/v))

onlineFold = subset(offlineSummary, posXY %in% permuteLocs[ , 1])

reshapeSS = function(data, varSignal = "signal", 
                     keepVars = c("posXY", "posX","posY"),
                     sampleAngle = FALSE, 
                     refs = seq(0, 315, by = 45)) {
  byLocation =
    with(data, by(data, list(posXY), 
                  function(x) {
                    if (sampleAngle) {
                      x = x[x$angle == sample(refs, size = 1), ]}
                    ans = x[1, keepVars]
                    avgSS = tapply(x[ , varSignal ], x$mac, mean)
                    y = matrix(avgSS, nrow = 1, ncol = 6,
                               dimnames = list(ans$posXY,
                                               names(avgSS)))
                    cbind(ans, y)
                  }))

  newDataSS = do.call("rbind", byLocation)
  return(newDataSS)
}

```
### Run Model

### Test
```{r}

offline = offline[ offline$mac != "00:0f:a3:39:dd:cd", ]

keepVars = c("posXY", "posX","posY", "orientation", "angle")

onlineCVSummary = reshapeSS(offline, keepVars = keepVars, 
                            sampleAngle = TRUE)

onlineFold = subset(onlineCVSummary, 
                    posXY %in% permuteLocs[ , 1])

offlineFold = subset(offlineSummary,
                     posXY %in% permuteLocs[ , -1])

estFold = predXY(newSignals = onlineFold[ , 6:11], 
                 newAngles = onlineFold[ , 4], 
                 offlineFold, numAngles = 3, k = 3)

actualFold = onlineFold[ , c("posX", "posY")]
calcError(estFold, actualFold)
```

### new PredXY
```{r}
predXY_new = function(newSignals, newAngles, trainData, 
                  numAngles = 1, k = 3){
  
  closeXY = list(length = nrow(newSignals))
  
  for (i in 1:nrow(newSignals)) {
    trainSS = selectTrain(newAngles[i], trainData, m = numAngles)
    closeXY[[i]] = findNN(newSignal = as.numeric(newSignals[i, ]),
                           trainSS)
  }

  estXY = lapply(closeXY, function(x)
                            sapply(x[ , 2:3], 
                                    function(x) mean(x[1:k])))
  estXY = do.call("rbind", estXY)
  return(estXY)
}
```

### Run K 1:20
```{r}
K = 20
err = rep(0, K)

for (j in 1:v) {
  message(j,'/',v,appendLF = F)
  onlineFold = subset(onlineCVSummary, 
                      posXY %in% permuteLocs[ , j])
  offlineFold = subset(offlineSummary,
                       posXY %in% permuteLocs[ , -j])
  actualFold = onlineFold[ , c("posX", "posY")]
  
  for (k in 1:K) {
    message('.',appendLF = F)
    estFold = predXY(newSignals = onlineFold[ , 6:11],
                     newAngles = onlineFold[ , 4], 
                     offlineFold, numAngles = 3, k = k)
    err[k] = err[k] + calcError(estFold, actualFold)
  }
  message('')
}
```


### Plotting
```{r}

plot(y = err, x = (1:K),  type = "l", lwd= 2,
     ylim = c(800, 2100),
     xlab = "Number of Neighbors",
     ylab = "Sum of Square Errors")

rmseMin = min(err)
kMin = which(err == rmseMin)[1]
segments(x0 = 0, x1 = kMin, y0 = rmseMin, col = gray(0.4), 
         lty = 2, lwd = 2)
segments(x0 = kMin, x1 = kMin, y0 = 1100,  y1 = rmseMin, 
         col = grey(0.4), lty = 2, lwd = 2)

#mtext(kMin, side = 1, line = 1, at = kMin, col = grey(0.4))
text(x = kMin - 2, y = rmseMin + 40, 
     label = as.character(round(rmseMin)), col = grey(0.4))

```

## Prediction
```{r}
estXYk5 = predXY(newSignals = onlineSummary[ , 6:11], 
                 newAngles = onlineSummary[ , 4], 
                 offlineSummary, numAngles = 3, k = 5)

calcError(estXYk5, actualXY)
```




